### 2016-09-18 appendix.tex
changed toolchain link

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages in its Linux distribution for performing
the commands, particularly the ones regarding the \texttt{util-linux} package
or if a command just fails.

>>>>

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages on his / her Linux distribution for performing
the commands, particularly the ones regarding the \texttt{util-linux} package
or if a command just fails.

### 2016-09-18 cross_compilation.tex
changed toolchain link

\subsection{Toolchain Setup}
To compile in a given architecture that is aimed for a different one,
the testbed administrator needs to install a toolchain in its \ac{PC}.
The toolchain is mandatory due to the different processor architectures
where the source can be compiled from.
Given that compiling a toolchaing can be an arduous task, we get the toolchain
recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
This toolchain is available from \cite{2016steinwurfrpitoolchain} and
it already contains the binaries for different compilers based on
\texttt{gcc 4.9}. In what follows, we extract the binaries adjusting
them to our coding style and compiling convention. For this, we use the
\texttt{\$\{TOOLCHAIN\}} directory as the working directory.
The testbed administrator may choose some other working directory of
its preference if desired. First, we create the toolchain directory:

@@

Later, we download a \ac{Raspi} toolchain with the binaries for a 64-bit Linux
distribution available in \cite{2016steinwurfrpitoolchain}. Finally, we unzip the
downloaded file. This is made as follows:

\begin{lstlisting}[]
$ wget http://files.steinwurf.com/toolchains/linux/raspberry-gcc-4.9.3/raspberry-gxx493-arm.zip
$ unzip raspberry-gxx493-arm.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

>>>>

\subsection{Toolchain Setup}
To compile in a given architecture that is aimed for a different one,
the testbed administrator needs to install a toolchain on his / her \ac{PC}.
The toolchain is mandatory due to the different processor architectures
where the source can be compiled from.
Given that compiling a toolchain can be an arduous task, we get the toolchain
recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
This toolchain is available from \cite{soerensen_chres_wiant_2016_154143} and
it already contains the binaries for different compilers based on
\texttt{gcc 4.9}. In what follows, we extract the binaries adjusting
them to our coding style and compiling convention. For this, we use the
\texttt{\$\{TOOLCHAIN\}} directory as the working directory.
The testbed administrator may choose some other working directory of
its preference if desired. First, we create the toolchain directory:

@@

Later, we download a \ac{Raspi} toolchain with the binaries for a 64-bit Linux
distribution available in \cite{soerensen_chres_wiant_2016_154143}. Finally, we unzip the
downloaded file. This is made as follows:

\begin{lstlisting}[]
$ wget https://zenodo.org/record/154328/files/raspberry-gxx493-arm.zip
$ unzip raspberry-gxx493-arm.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

### 2016-09-18 image_setup.tex
changed toolchain link

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are also available there. The testbed
administrator has to move these files to his / her own \ac{HTTP} server.
To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

>>>>

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are also available there. The testbed
administrator has to move these files to his / her own \ac{HTTP} server.
To get started, the testbed administrator must open a Linux
shell (terminal) on his / her \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

### 2016-09-18 overlay_fs.tex
Corrections
changed toolchain link

We create a partition for the home directory after the root partition.
To do this, we again use \texttt{fdisk} to find the next available sector
in the image. To verify the new available space for the full image and observe
the next available sector, we run:
\begin{lstlisting}[]
$ fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

We notice that one GB is now available to be used in the partitions. Also, we
observe the new partition should start at sector 4806656. To create it, we
show the introduced command:% and its output:

%\begin{lstlisting}[]
%$ (echo n; echo p; echo 3; echo 4806656; echo ; echo w) | fdisk ${IMAGE}.img
%
%Command (m for help): Partition type:
%   p primary (1 primary, 0 extended, 3 free)
%   e extended
%Partition number (1-4, default 3): Using default value 3
%First sector (2048-6903807, default 2048): Last sector, +sectors or
%+size{K,M,G} (4806656-6903807, default 6903807): Using default value 6903807
%
%Command (m for help): The partition table has been altered!
%
%Syncing disks.
%\end{lstlisting}
\begin{lstlisting}[]
$ fdisk ${IMAGE}.img <<EOF
n
p
3
4806656

@@

%\textcolor{red}{SHOW NEW PARTITION TABLE HERE}

@@

\subsection{Writing Customized Image to SD Memory Cards}
For a basic system setup, the final step is to write the customized image
to all the memory cards before they can be used in the \ac{Raspi}s. For
our current considered system, we do this manually for each card.
The testbed administrator needs to insert each memory card in its \ac{PC}
and follow the procedure in this section. A given card will be available as
\texttt{/dev/mmcblkX} or \texttt{/dev/sdYX} where \texttt{X} is a natural number
and \texttt{Y} is a letter.

>>>>

We create a partition for the home directory after the root partition.
To do this, we again use \texttt{fdisk} to find the next available sector
in the image. To verify the new available space for the full image and observe
the next available sector, we run:
\begin{lstlisting}[]
$ fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

We notice that one GB is now available to be used in the partitions. Also, we
observe the new partition should start at sector 4806656. To create it, we
use \texttt{fdisk} as follows:% and its output:

%\begin{lstlisting}[]
%$ (echo n; echo p; echo 3; echo 4806656; echo ; echo w) | fdisk ${IMAGE}.img
%
%Command (m for help): Partition type:
%   p primary (1 primary, 0 extended, 3 free)
%   e extended
%Partition number (1-4, default 3): Using default value 3
%First sector (2048-6903807, default 2048): Last sector, +sectors or
%+size{K,M,G} (4806656-6903807, default 6903807): Using default value 6903807
%
%Command (m for help): The partition table has been altered!
%
%Syncing disks.
%\end{lstlisting}
\begin{lstlisting}[]
$ fdisk ${IMAGE}.img << EOF
n
p
3
4806656

@@

\textcolor{red}{SHOW NEW PARTITION TABLE HERE}

@@

\subsection{Writing Customized Image to SD Memory Cards}
For a basic system setup, the final step is to write the customized image
to all the memory cards before they can be used in the \ac{Raspi}s. For
our current considered system, we do this manually for each card.
The testbed administrator needs to insert each memory card in his / her \ac{PC}
and follow the procedure in this section. A given card will be available as
\texttt{/dev/mmcblkX} or \texttt{/dev/sdYX} where \texttt{X} is a natural number
and \texttt{Y} is a letter.
